---
title: "IODS course final project"
author: "Volkova Anastasia"
date: "18.12.2017"
e-mail: "mibisako@gmail.com" 
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 2
    fig_caption: true
    fig_width: 6
    fig_height: 4
    code_folding: hide
---

##**IODS Final Project**

*I want to explore deep and surface learning approaches using binary logistic regression. I suppose that older students prefer deep LA in contrast to younger. I do not expect any gender differences. I also suppose that general attitude towards statistics may highly influence the LA - if people are not interested, they are more likely to use surface approach.*

[Here you can find a link to my data wrangling script.](https://www.google.com "Wrangling learning2014")

```{r}
library(ggplot2)
library(GGally)
library(dplyr)
library(tidyr)
task<-read.table("C:/Users/Missao/Documents/GitHub/IODS-final/learning2014.txt", header=TRUE, sep="\t")
str(task)
dim(task)
names(task)
```

The dataset is the result of a survey on teaching and learning conducted by Kimmo Vehkalahti in 2014. Additional information about this research project can be found here https://www.slideshare.net/kimmovehkalahti/the-relationship-between-learning-approaches-and-students-achievements-in-an-introductory-statistics-course-in-finland and there http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS2-meta.txt.


It has 166 observations of 11 variables. Learning approaches variables were created out of related questions and then adjusted.


We are most interested in two approaches, that lies in the DS variable, where 


**1** = Deep: Seeking Meaning, Relating Ideas, Use of Evidence, and 


**0** = Surface: Lack of Purpose, Unrelated Memorising, Syllabus-boundness.




```{r}
library(ggplot2)
library(GGally)
ggpairs(task, lower = list(combo = wrap("facethist", bins = 20)))
```

There were more females than males in the study. The age of people ranges from 17 to 55 years, and the distribution is positively skewed. Three learning approaches (deep, surface and strategic) have more or less close to normal distributions, though "deep approach" is left-skewed. One of the highest correlation is between deep and surface approaches, and as logically expected, it is negative (those who have high scores on questions related to deep approach, have low scores on surface questions). It seems that the strategic approach is most related to the deep approach (positive correlation). 

To see what can influence the probabilities of choosing deep or surface approach a *binary logistic regression* will be used. 



```{r}
m1 <- glm(DS ~ strategic_adjusted + gender + Age, data = task, family = "binomial")
summary(m1)
OR <- coef(m1) %>% exp # compute odds ratios (OR)
CI <- confint(m1) %>% exp # compute confidence intervals (CI)
cbind(OR, CI)
```

According to R-squared, the final model explains around 18,5% of variance in the outcome. As it seems from the p-values, the age and gender are statistically insignificant in relation to the choosing of learning approaches. However, as expected, those people who have high scores on strategic learning approach questions, have more chances to use deep learning approach. The better is the general attitude towards statistics, the better the exam scores are (3,5255 score per each unit of attitude increase). 

